---
title: "Course Project 1"
author: "Martin Mason"
date: "9 August 2016"
output: html_document
---

####Loading the Document into R

First we download the data the 
```{r eval=FALSE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activityData.zip", method="curl")
```

Then we unzip the data
```{r}
unzip("activityData.zip")
```

Then we store data in a variable called activity
```{r}
myCSV <- read.csv("activity.csv", header = TRUE)
```

##What is the mean total number of steps taken per day?
First we create an array called stepsPerDay with the number of steps for each day
```{r}
stepsPerDay <- tapply(myCSV$steps, myCSV$date, sum, na.rm=TRUE)
```

Which give us the following array
```{r echo=FALSE}
stepsPerDay
```

This can in turn be represented as a histogram
```{r}
hist(stepsPerDay)
```

The mean and median can be calculate in the followin way
```{r results="hide"}
myMedian <- median(stepsPerDay)
myMean <- mean(stepsPerDay)
```
This gives us a mean of `r myMean` and a median of `r myMedian`.

##What is the average daily activity pattern?

Here is the time-series plot:

```{r echo = FALSE}
plot(unique(myCSV$interval), tapply(myCSV$steps, myCSV$interval, mean, na.rm=TRUE), type="l")
```

I admit I was pretty unsure of this section but
the calculation max of the code

```{r}
myData <- tapply(myCSV$steps, myCSV$interval, mean, na.rm=TRUE)
```

To find the 5 minute interval which was highest I did the following:

```{r}
highestInterval <-  names(myData[myData == max(myData)])
```

Returns the value `r max(myData)`, showing that it was in fact interval number `r highestInterval`.

##Inputing missing values

First we count the number of rows with NAs:

```{r}
NAs <- (!complete.cases(myCSV))
NAs <- NAs[NAs == TRUE]
NACount <- length(NAs)
return(NACount)
```

This means that the number of rows that are incomplete, e.g. contain NAs, is `r NACount`.




